`You Draw It': Implementation of visually fitted trends with \texttt{r2d3}

### 2022-11-14 2022-sdss-you-draw-it-manuscript.qmd
edited discussion

The 'You Draw It' graphical task presented in this work is an example of an interactive graphic. 
While any communication between user and device can be considered an interaction, interactive graphics should be defined as a user-driven direct manipulation of plots and plotting elements with an immediate reaction (Unwin 1999, Becker et al. 1987, Eick & Wills 1995).
Here we discuss the use of 'You Draw It' as a tool for interactive reading, interactive testing, and interactive teaching as a way to increase user experience and user engagement with data.
The New York Times first introduced the 'You Draw It' feature as a tool for interactive reading.
The inclusion of interactive graphics in news and media have been used to inform readers in efforts to reduce misconceptions and aid in storytelling [@geidner2015role; @george2011interactive; @george2013storytelling].
In our work, we demonstrated the use of the 'You Draw It' method as a tool for interactive testing of graphics.
We collect data of the user drawn lines and compare to statistical results in order to understand perceptual biases.
While this work has many possible directions for future work, there is a strong application to extend the 'You Draw It' tool for interactive teaching.
The line of best fit is introduced in secondary schools as a student's first encounter with the concept of statistical association [@casey2016students].
@casey2015examining and @casey2016finding conducted task based interviews to examine student's conceptions of the line of best fit prior to introducing the topic.
The implications of using interactive teaching tools for the line of best fit can help students make connections between the shift in slope and the location of the best fit line [@nagle2017slope].
@casey2015teachers conducted similar task based interviews wish pre-service mathematics teachers in order to understand the implications regarding teacher preparation for the teaching of this topic. 
This demonstrates the use of the 'You Draw It' interactive tool for training purposes both within a classroom setting and for those in the workforce.

>>>>

The 'You Draw It' graphical task presented in this work is an example of an interactive graphic; an interactive graphic is defined as a user-driven direct manipulation of the plot followed by an immediate reaction [@unwin1999requirements; @eick1995high; @becker1987dynamic].
Here we discuss the use of 'You Draw It' as a tool for interactive reading, interactive testing, and interactive teaching as a way to increase user experience and user engagement with data.
The New York Times first introduced the 'You Draw It' feature as a tool for interactive reading.
The inclusion of interactive graphics in news and media have been used to inform readers in efforts to reduce misconceptions and aid in storytelling [@geidner2015role; @george2011interactive; @george2013storytelling].
In our work, we demonstrated the use of the 'You Draw It' method as a tool for interactive testing of graphics.
We collected and compared data from user drawn lines to statistical results in order to understand perceptual biases.
While this work has many possible directions for future work, there is a strong application to extend the 'You Draw It' tool for interactive teaching.
The line of best fit is introduced in secondary schools as a student's first encounter with the concept of statistical association between multiple variables [@casey2016students].
@casey2015examining and @casey2016finding conducted task based interviews to examine student's conceptions of the line of best fit prior to introducing the topic.
The implications of using interactive teaching tools for the line of best fit can help students make connections between the shift in slope and the location of the best fit line [@nagle2017slope].
@casey2015teachers conducted similar task based interviews wish pre-service mathematics teachers in order to understand the effects regarding teacher preparation for the teaching of this topic. 
This demonstrates the use of the 'You Draw It' interactive tool for training purposes both within a classroom setting and for those in the workforce.

### 2022-11-14 2022-sdss-you-draw-it-manuscript.tex
edited discussion

\title{`You Draw It': Implementation of visually fitted trends with
\texttt{r2d3}}
\author{}
\date{}

\begin{document}
\maketitle
\begin{abstract}
How do statistical regression results compare to intuitive, visually
fitted results? Fitting lines by eye through a set of points has been
explored since the 20th century. Common methods of fitting trends by eye
involve maneuvering a string, black thread, or ruler until the fit is
suitable, then drawing the line through the set of points. In 2015, the
New York Times introduced an interactive feature, called `You Draw It,'
where readers are asked to input their own assumptions about various
metrics and compare how these assumptions relate to reality. This
research is intended to implement `You Draw It', adapted from the New
York Times, as a way to measure the patterns we see in data. In this
paper, we describe the adaptation of an old tool for graphical testing
and evaluation, eye-fitting, for use in modern web-applications suitable
for testing statistical graphics. We present an empirical evaluation of
this testing method for linear regression, and briefly discuss an
extension of this method to non-linear applications.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, enhanced, interior hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}, frame hidden, boxrule=0pt]}{\end{tcolorbox}}\fi

@@

Initial studies in the 20th century explored the use of fitting lines by
eye through a set of points (Finney 1951; Mosteller et al. 1981; Unwin
and Wills 1988). Common methods of fitting trends by eye involved
maneuvering a string, black thread, or ruler until the fit was suitable,
then drawing the line through the set of points. Results from these
early studies provided groundwork for visually judging slopes and
selecting accurate starting values for common iterative calculations.
Recently, Ciccione and Dehaene (2021) conducted a comprehensive set of
studies investigating human ability to detect trends in graphical
representations from a psychophysical approach. This set of studies
asked participants to judge trends, estimate slopes, and conduct
extrapolation by using a track-pad to adjust the tilt of a line on the
screen. Results indicated the slopes participants reported were always
in excess of the ideal slopes, both in the positive and in the negative
direction, and those biases increased with noise and with number of
points. This supported the results found in Mosteller et al. (1981) and
suggested that participants might use Deming regression (Deming 1943),
which minimizes the Euclidean distance of points from the line, when
fitting a line to a noisy scatterplot.

@@

The `You Draw It' graphical task presented in this work is an example of
an interactive graphic. While any communication between user and device
can be considered an interaction, interactive graphics should be defined
as a user-driven direct manipulation of plots and plotting elements with
an immediate reaction (Unwin 1999, Becker et al.~1987, Eick \& Wills
1995). Here we discuss the use of `You Draw It' as a tool for
interactive reading, interactive testing, and interactive teaching as a
way to increase user experience and user engagement with data. The New
York Times first introduced the `You Draw It' feature as a tool for
interactive reading. The inclusion of interactive graphics in news and
media have been used to inform readers in efforts to reduce
misconceptions and aid in storytelling (Geidner et al. 2015;
George-Palilonis and Spillman 2011, 2013). In our work, we demonstrated
the use of the `You Draw It' method as a tool for interactive testing of
graphics. We collect data of the user drawn lines and compare to
statistical results in order to understand perceptual biases. While this
work has many possible directions for future work, there is a strong
application to extend the `You Draw It' tool for interactive teaching.
The line of best fit is introduced in secondary schools as a student's
first encounter with the concept of statistical association (Casey and
Nagle 2016). Casey (2015) and Casey (2016) conducted task based
interviews to examine student's conceptions of the line of best fit
prior to introducing the topic. The implications of using interactive
teaching tools for the line of best fit can help students make
connections between the shift in slope and the location of the best fit
line (Nagle, Casey, and Moore-Russo 2017). Casey and Wasserman (2015)
conducted similar task based interviews wish pre-service mathematics
teachers in order to understand the implications regarding teacher
preparation for the teaching of this topic. This demonstrates the use of
the `You Draw It' interactive tool for training purposes both within a
classroom setting and for those in the workforce.

@@

@@

@@

>>>>

\title{`You Draw It': Implementation of visually fitted trends with
\texttt{r2d3}}
% \author{}
% \date{}

\begin{document}
\maketitle
\begin{abstract}
How do statistical regression results compare to intuitive, visually
fitted results? Fitting lines by eye through a set of points has been
explored since the 20th century. Common methods of fitting trends by eye
involve maneuvering a string, black thread, or ruler until the fit is
suitable, then drawing the line through the set of points. In 2015, the
New York Times introduced an interactive feature, called `You Draw It,'
where readers are asked to input their own assumptions about various
metrics and compare how these assumptions relate to reality. This
research is intended to implement `You Draw It', adapted from the New
York Times, as a way to measure the patterns we see in data. In this
paper, we describe the adaptation of an old tool for graphical testing
and evaluation, eye-fitting, for use in modern web-applications suitable
for testing statistical graphics. We present an empirical evaluation of
this testing method for linear regression, and briefly discuss an
extension of this method to non-linear applications.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, sharp corners, borderline west={3pt}{0pt}{shadecolor}, breakable, enhanced, frame hidden, interior hidden]}{\end{tcolorbox}}\fi

@@

Initial studies in the 20th century explored the use of fitting lines by
eye through a set of points (Finney 1951; Mosteller et al. 1981; A. R.
Unwin and Wills 1988). Common methods of fitting trends by eye involved
maneuvering a string, black thread, or ruler until the fit was suitable,
then drawing the line through the set of points. Results from these
early studies provided groundwork for visually judging slopes and
selecting accurate starting values for common iterative calculations.
Recently, Ciccione and Dehaene (2021) conducted a comprehensive set of
studies investigating human ability to detect trends in graphical
representations from a psychophysical approach. This set of studies
asked participants to judge trends, estimate slopes, and conduct
extrapolation by using a track-pad to adjust the tilt of a line on the
screen. Results indicated the slopes participants reported were always
in excess of the ideal slopes, both in the positive and in the negative
direction, and those biases increased with noise and with number of
points. This supported the results found in Mosteller et al. (1981) and
suggested that participants might use Deming regression (Deming 1943),
which minimizes the Euclidean distance of points from the line, when
fitting a line to a noisy scatterplot.

@@

The `You Draw It' graphical task presented in this work is an example of
an interactive graphic; an interactive graphic is defined as a
user-driven direct manipulation of the plot followed by an immediate
reaction (A. Unwin 1999; Eick and Wills 1995; Becker, Cleveland, and
Wilks 1987). Here we discuss the use of `You Draw It' as a tool for
interactive reading, interactive testing, and interactive teaching as a
way to increase user experience and user engagement with data. The New
York Times first introduced the `You Draw It' feature as a tool for
interactive reading. The inclusion of interactive graphics in news and
media have been used to inform readers in efforts to reduce
misconceptions and aid in storytelling (Geidner et al. 2015;
George-Palilonis and Spillman 2011, 2013). In our work, we demonstrated
the use of the `You Draw It' method as a tool for interactive testing of
graphics. We collected and compared data from user drawn lines to
statistical results in order to understand perceptual biases. While this
work has many possible directions for future work, there is a strong
application to extend the `You Draw It' tool for interactive teaching.
The line of best fit is introduced in secondary schools as a student's
first encounter with the concept of statistical association between
multiple variables (Casey and Nagle 2016). Casey (2015) and Casey (2016)
conducted task based interviews to examine student's conceptions of the
line of best fit prior to introducing the topic. The implications of
using interactive teaching tools for the line of best fit can help
students make connections between the shift in slope and the location of
the best fit line (Nagle, Casey, and Moore-Russo 2017). Casey and
Wasserman (2015) conducted similar task based interviews wish
pre-service mathematics teachers in order to understand the effects
regarding teacher preparation for the teaching of this topic. This
demonstrates the use of the `You Draw It' interactive tool for training
purposes both within a classroom setting and for those in the workforce.

@@

\leavevmode\vadjust pre{\hypertarget{ref-becker1987dynamic}{}}%
Becker, Richard A, William S Cleveland, and Allan R Wilks. 1987.
{``Dynamic Graphics for Data Analysis.''} \emph{Statistical Science} 2
(4): 355--83.

@@

\leavevmode\vadjust pre{\hypertarget{ref-eick1995high}{}}%
Eick, Stephen G, and Graham J Wills. 1995. {``High Interaction
Graphics.''} \emph{European Journal of Operational Research} 81 (3):
445--59.

@@

\leavevmode\vadjust pre{\hypertarget{ref-unwin1999requirements}{}}%
Unwin, Antony. 1999. {``Requirements for Interactive Graphics Software
for Exploratory Data Analysis.''} \emph{Computational Statistics} 14
(1): 7--22.
